Military mechanization as a complex adaptive system

### 2018-07-19 paper2/paper2.tex
more wek

Complex systems theory \citep[ultimately stemming to][]{simon1996artificial}
offers a way forward. Although we cannot separate the many parts from the whole,
we would like to modify the value at one of the nodes and observe the resultant system state. This sort of
experimentation might allow us observe the system's sensitivity (or lack thereof) to changes
in input. But, as previously covered, one generally cannot conduct
experimental trials in international relations.  A modeling paradigm that would let us do so is needed.

@@

There is one agent type: states. Agents exist at a fixed location and do not
move. The set of agents encompasses all diplomatically-recognized 
states in the world at a particular point in time, minus microstates (defined as
those with a population less than 750,000). Some other states were excluded; for
why, see the \textbf{Time} section. Figure 1 depicts the set of agents.

@@

An important note is that these three factors---neighbors, enemies, and
defeat in counter-insurgency---correspond to the factors identified by 
\citet{sechser2010army} as most important for mechanization. Because the research
goal is essentially to see whether their findings stand up to agentization,
it is critical to make agent cognition work in about the way that their findings
predict. Inclusion of other factors not identified as important for
mechanization---e.g., alliances, economic factors,
culture, etc.---would be interesting but counter-productive to the research
question. They are discussed along with this model's findings.

@@

@@

@@

\printbibliography[heading=bibnumbered]

In summary, the orthodox view of mechanization is that it is a
reflexive process: states choose their mechanization posture largely based on
that neighbors, enemies, and recent military experiences (e.g. defeat in
counter-insurgency). This paper argues
that this view---based on traditional longitudinal data analysis---is ripe for
agentization. It implicitly argues for autonomous agents (states) that have
cognition regarding mechanization that takes into account other agents (e.g.
enemies), spatiality (e.g. neighbors), and learning from experience. This paper
builds such a model and validates it against the empirical record, using the
National Mechanization Index of \citet{sechser2010army}. It finds that the 
orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
causes states to
mechanize much faster and to a greater degree than occurs in the real-world.
This implies the existence of some countervailing factor (possibly a negative
feedback loop) that mitigates the extent of mechanization. Candidates for what
this could be mostly center on ``internal'' factors---e.g., culture, economy,
and so on---as opposed to the ``external'', security environment-centric
orthodox view.

are ripe for agentization. The
conclusions drawn implicitly claim
that states are autonomous agents; that they have
cognition regarding how to perceive the outside world
and how to react to it; that spatiality and
neighborhood matter; that inter-agent relationships matter; and that
agents learn from experience. Agent-based modeling is uniquely well-suited to
represent such phenomena .

Importantly, the findings of \citet{sechser2010army} imply a positive feedback
loop. State A has high
mechanization, causing enemy State B to raise their mechanization, causing State
A to raise their mechanization... and so on. Such feedback loops are a hallmark
of complex systems and again are well-suited for representation by agent-based
modeling. This specific feedback loop is often called
the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
of agent-based modeling \citep{schelling1960strategy,schelling2006micromotives}.

\begin{table}[h]
	\centering
	\begin{tabular}{l r}
		\hline
		\textbf{Variable} & \textbf{VIF} \\ 
		\hline
		log\_mech\_enemies & 8.76 \\ 
		enemies\_dummy & 9.12 \\ 
		log\_mech\_neighbors & 2.19 \\ 
		neighbors\_dummy & 2.03 \\ 
		log\_mountainous & 1.26 \\ 
		antimech\_lesson\_2 & 1.21 \\ 
		promech\_lesson\_2 & 1.11 \\ 
		democracy & 1.69 \\ 
		military\_govt & 1.13 \\ 
		instability & 1.12 \\ 
		log\_gdp\_cap & 2.10 \\ 
		iron\_steel & 1.32 \\ 
		log\_mech\_allies & 12.64 \\ 
		allies\_dummy & 12.59 \\ 
		cold\_war\_bloc & 1.48 \\ 
		igos & 2.03 \\ 
		post\_soviet & 1.33 \\ 
		log\_mech\_lag2 & 1.57 \\ 
		\hline
	\end{tabular}
\end{table}
\end{document}

>>>>

Complex systems theory \citep[e.g.][]{simon1996sciences}
offers a way forward. Although we cannot separate the many parts from the whole,
we would like to modify the value at one of the nodes and observe the resultant system state. This sort of
experimentation might allow us observe the system's sensitivity (or lack thereof) to changes
in input. But, as previously covered, one generally cannot conduct
experimental trials in international relations.  A modeling paradigm that would let us do so is needed.

@@

There is one agent type: states. Agents exist at a fixed location and do not
move. The set of agents encompasses all diplomatically-recognized 
states in the world at a particular point in time, minus microstates (defined as
those with a population less than 750,000). Some other states were excluded; for
why, see the \textbf{Time} section. Figure 3 depicts the set of agents.

@@

An important note is that these three factors---neighbors, enemies, and
defeat in counter-insurgency---correspond to the factors identified by 
\citet{sechser2010army} as most important for mechanization. Because the research
goal is essentially to see whether their findings stand up to agentization,
it is critical to make agent cognition work in about the way that their findings
predict. Inclusion of other factors not identified as important for
mechanization---e.g., alliances, culture, etc.---would be interesting 
but counter-productive to the research
question. They are discussed along with this model's findings.

@@

BEEF UP THIS SECTION

@@

\subsection{Future Research}

@@

\printbibliography[heading=bibnumbered]
\newpage
\appendix
\centering
\section*{Appendix A}
\begin{table}[h]
	\centering
	\caption{Variance inflation factor, \citet{sechser2010army}}
	\begin{tabular}{l r}
		\hline
		\textbf{Variable} & \textbf{VIF} \\ 
		\hline
		log\_mech\_enemies & 8.76 \\ 
		enemies\_dummy & 9.12 \\ 
		log\_mech\_neighbors & 2.19 \\ 
		neighbors\_dummy & 2.03 \\ 
		log\_mountainous & 1.26 \\ 
		antimech\_lesson\_2 & 1.21 \\ 
		promech\_lesson\_2 & 1.11 \\ 
		democracy & 1.69 \\ 
		military\_govt & 1.13 \\ 
		instability & 1.12 \\ 
		log\_gdp\_cap & 2.10 \\ 
		iron\_steel & 1.32 \\ 
		log\_mech\_allies & 12.64 \\ 
		allies\_dummy & 12.59 \\ 
		cold\_war\_bloc & 1.48 \\ 
		igos & 2.03 \\ 
		post\_soviet & 1.33 \\ 
		log\_mech\_lag2 & 1.57 \\ 
		\hline
	\end{tabular}
\end{table}
\end{document}
